---
title: "Pitcher Precision"
author: "Sasank Vishnubhatla"
date: "4/17/2019"
runtime: shiny
output: html_document
---

```{r setup, include=FALSE, results='hide'}
knitr::opts_chunk$set(echo = TRUE)
```

Last Update: `r Sys.time()`

# Libraries

Let's load some libraries in first.

```{r libs}
library(baseballr)
library(pitchRx)
library(tidyverse)
```

Let's also clean out environment.

```{r clean}
rm(list = ls())
```

With these libraries, we can get out data as well as visaulize it. Let's take a look at some players to see what we can look at.

# Data Loading

Here are the list of players I will be looking at.

* Noah Syndergaard - Player ID: 592789
* Patrick Corbin - Player ID: 571578
* Felipe Vazquez - Player ID: 553878
* Marcus Stroman - Player ID: 573186
* Justin Verlander - Player ID: 434378
* Blake Treinen - Player ID: 595014

Let's now scrape the data for each player.

```{r disco_data, message = FALSE}
scrape.data = function(start, id) {
    data = scrape_statcast_savant(start_date = start,
                                  end_date = format(Sys.time(), "%Y-%m-%d"),
                                  playerid = id,
                                  player_type = 'pitcher')
    data
}

start = "2019-01-01"

syndergaard.data = scrape.data(start, 592789)
corbin.data = scrape.data(start, 571578)
vazquez.data = scrape.data(start, 553878)
stroman.data = scrape.data(start, 573186)
verlander.data = scrape.data(start, 434378)
treinen.data = scrape.data(start, 595014)
```

Now with our data, let's get the information we want out of it.

```{r filter}
filter.data = function(data) {
    filtered = data.frame(name = data %>% pull(player_name),
                          pitch = data %>% pull(pitch_type),
                          outcome = data %>% pull(type),
                          date = data %>% pull(game_date),
                          event = data %>% pull(events),
                          descrip = data %>% pull(description),
                          xcoord = data %>% pull(plate_x),
                          ycoord = data %>% pull(plate_z),
                          xmove = data %>% pull(pfx_x),
                          ymove = data %>% pull(pfx_z),
                          velo = data %>% pull(effective_speed),
                          spin = data %>% pull(release_spin_rate),
                          exvelo = data %>% pull(launch_speed),
                          exang = data %>% pull(launch_angle),
                          contact = data %>% pull(launch_speed_angle),
                          year = substring(data %>% pull(game_date), 0, 4))
    filtered$exvelo[is.na(filtered$exvelo)] = 0
    filtered$exang[is.na(filtered$exang)] = 0
    filtered$contact[is.na(filtered$contact)] = 0
    filtered
}

syndergaard = filter.data(syndergaard.data)
corbin = filter.data(corbin.data)
stroman = filter.data(stroman.data)
treinen = filter.data(treinen.data)
vazquez = filter.data(vazquez.data)
verlander = filter.data(verlander.data)
```

With this filtered data, we have selected the following columns:

* Name
* Pitch type
* Pitch outcome (Strike, Ball, In play)
* Game Date
* Event
* Description of event
* X coordinate (horizontal location) of pitch
* Y coordinate (vertical location) of pitch
* Horizontal movement (X coordinate movement) of pitch
* Vertical movement (Y coordinate movement) of pitch
* Velocity of pitch
* Spin rate of pitch
* Exit velocity of pitch
* Launch angle of pitch

## Visualization {.tabset .tabset-fade}

Let's start visualizing some of this data. Before that, let me define a strikezone. This strikezone was taken from the website [Baseball with R](https://baseballwithr.wordpress.com/2013/12/12/plotting-pitches-ggplot2-tips-and-tricks/)

```{r strikezone}
topKzone = 3.5
botKzone = 1.6
inKzone = -.95
outKzone = 0.95
kZone = data.frame(x = c(inKzone, inKzone, outKzone, outKzone, inKzone),
                   y = c(botKzone, topKzone, topKzone, botKzone, botKzone))
```

### Location via Outcome {.tabset .tabset-fade}

Let's look at pitch location with if the pitch is a ball or strike. We know X is hit into play, B is ball, and S is any type of strike.

```{r pitch_heatmap_out}
graph.pitch.heatmap.out = function(player) {
    graph = ggplot(player) +
        geom_jitter(aes(x = player$xcoord,
                        y = player$ycoord,
                        color = player$outcome)) +
        xlab("Horizontal Position") +
        ylab("Vertical Position") +
        ggtitle(paste(player$name[1], player$year[1], "Outcome", sep = " ")) +
        labs(color = "Pitch Outcome") +
        theme_minimal() + geom_path(aes(x, y), data = kZone)
    graph
}
```

#### Patrick Corbin

```{r}
corbin.heatmap.out = graph.pitch.heatmap.out(corbin)
corbin.heatmap.out
```

#### Marcus Stroman

```{r}
stroman.heatmap.out = graph.pitch.heatmap.out(stroman)
stroman.heatmap.out
```

#### Noah Syndergaard

```{r}
syndergaard.heatmap.out = graph.pitch.heatmap.out(syndergaard)
syndergaard.heatmap.out
```

#### Blake Treinen

```{r}
treinen.heatmap.out = graph.pitch.heatmap.out(treinen)
treinen.heatmap.out
```

#### Felipe Vazquez

```{r}
vazquez.heatmap.out = graph.pitch.heatmap.out(vazquez)
vazquez.heatmap.out
```

#### Justin Verlander

```{r}
verlander.heatmap.out = graph.pitch.heatmap.out(verlander)
verlander.heatmap.out
```

### Location via Type {.tabset .tabset-fade}

Let's look at pitch location via pitch type.

```{r pitch_heatmap_type}
graph.pitch.heatmap.type = function(player) {
    graph = ggplot(player) +
        geom_jitter(aes(x = player$xcoord,
                        y = player$ycoord,
                        color = player$pitch)) +
        xlab("Horizontal Position") +
        ylab("Vertical Position") +
        ggtitle(paste(player$name[1], player$year[1], "Type", sep = " ")) +
        labs(color = "Pitch Type") +
        theme_minimal() + geom_path(aes(x, y), data = kZone)
    graph
}
```

#### Patrick Corbin

```{r}
corbin.heatmap.type = graph.pitch.heatmap.type(corbin)
corbin.heatmap.type
```

#### Marcus Stroman

```{r}
stroman.heatmap.type = graph.pitch.heatmap.type(stroman)
stroman.heatmap.type
```

#### Noah Syndergaard

```{r}
syndergaard.heatmap.type = graph.pitch.heatmap.type(syndergaard)
syndergaard.heatmap.type
```

#### Blake Treinen

```{r}
treinen.heatmap.type = graph.pitch.heatmap.type(treinen)
treinen.heatmap.type
```

#### Felipe Vazquez

```{r}
vazquez.heatmap.type = graph.pitch.heatmap.type(vazquez)
vazquez.heatmap.type
```

#### Justin Verlander

```{r}
verlander.heatmap.type = graph.pitch.heatmap.type(verlander)
verlander.heatmap.type
```

### Location via Velocity {.tabset .tabset-fade}

Let's look at pitch location via velocity.

```{r pitch_heatmap_velo}
graph.pitch.heatmap.velo = function(player) {
    graph = ggplot(player) +
        geom_jitter(aes(x = player$xcoord,
                        y = player$ycoord,
                        color = player$velo)) +
        xlab("Horizontal Position") +
        ylab("Vertical Position") +
        ggtitle(paste(player$name[1], player$year[1], "Velocity", sep = " ")) +
        labs(color = "Velocity") +
        scale_color_gradient(low = "blue", high = "red") +
        theme_minimal() + geom_path(aes(x, y), data = kZone)
    graph
}
```

#### Patrick Corbin

```{r}
corbin.heatmap.velo = graph.pitch.heatmap.velo(corbin)
corbin.heatmap.velo
```

#### Marcus Stroman

```{r}
stroman.heatmap.velo = graph.pitch.heatmap.velo(stroman)
stroman.heatmap.velo
```

#### Noah Syndergaard

```{r}
syndergaard.heatmap.velo = graph.pitch.heatmap.velo(syndergaard)
syndergaard.heatmap.velo
```

#### Blake Treinen

```{r}
treinen.heatmap.velo = graph.pitch.heatmap.velo(treinen)
treinen.heatmap.velo
```

#### Felipe Vazquez

```{r}
vazquez.heatmap.velo = graph.pitch.heatmap.velo(vazquez)
vazquez.heatmap.velo
```

#### Justin Verlander

```{r}
verlander.heatmap.velo = graph.pitch.heatmap.velo(verlander)
verlander.heatmap.velo
```

### Movement {.tabset .tabset-fade}

To view the movement, let's just determine the average movement for each type of pitch that each player has. First let's make a few helpful functions for us.

```{r pitch_mov}
graph.pitch.xmovement = function(player) {
    graph = ggplot(player) +
        geom_boxplot(aes(x = player$pitch,
                         y = player$xmove,
                         color = player$pitch)) +
        coord_flip() +
        labs(color = "Pitch Type") +
        xlab("Pitch Type") + ylab("Horizontal Movement") +
        ggtitle(paste(player$name[1], player$year[1], "Horizontal Movement", sep = " ")) +
        theme_minimal()
}

graph.pitch.ymovement = function(player) {
    graph = ggplot(player) +
        geom_boxplot(aes(x = player$pitch,
                         y = player$ymove,
                         color = player$pitch)) +
        labs(color = "Pitch Type") +
        xlab("Pitch Type") + ylab("Vertical Movement") +
        ggtitle(paste(player$name[1], player$year[1], "Vertical Movement", sep = " ")) +
        theme_minimal()
}
```

#### Patrick Corbin

```{r}
corbin.xmove = graph.pitch.xmovement(corbin)
corbin.ymove = graph.pitch.ymovement(corbin)
corbin.xmove
corbin.ymove
```

#### Marcus Stroman

```{r}
stroman.xmove = graph.pitch.xmovement(stroman)
stroman.ymove = graph.pitch.ymovement(stroman)
stroman.xmove
stroman.ymove
```

#### Noah Syndergaard

```{r}
syndergaard.xmove = graph.pitch.xmovement(syndergaard)
syndergaard.ymove = graph.pitch.ymovement(syndergaard)
syndergaard.xmove
syndergaard.ymove
```

#### Blake Treinen

```{r}
treinen.xmove = graph.pitch.xmovement(treinen)
treinen.ymove = graph.pitch.ymovement(treinen)
treinen.xmove
treinen.ymove
```

#### Felipe Vazquez

```{r}
vazquez.xmove = graph.pitch.xmovement(vazquez)
vazquez.ymove = graph.pitch.ymovement(vazquez)
vazquez.xmove
vazquez.ymove
```

#### Justin Verlander

```{r}
verlander.xmove = graph.pitch.xmovement(verlander)
verlander.ymove = graph.pitch.ymovement(verlander)
verlander.xmove
verlander.ymove
```

### Velocity {.tabset .tabset-fade}

We need to separate each pitch first by type. Then we can see how the pitch's velocity changed over time.

```{r pitch_velo}
graph.pitch.velo = function(player) {
    graph = ggplot(player) +
        geom_line(aes(x = 1:length(player$velo),
                      y = player$velo,
                      color = player$pitch)) +
        xlab("Pitches Thrown") + ylab("Velocity") + labs(color = "Pitch Type") +
        ggtitle(paste(player$name[1], player$year[1], "Velocity Chart", sep = " ")) +
        theme_minimal()
}
```

#### Patrick Corbin

```{r}
corbin.velo = graph.pitch.velo(corbin)
corbin.velo
```

#### Marcus Stroman

```{r}
stroman.velo = graph.pitch.velo(stroman)
stroman.velo
```

#### Noah Syndergaard

```{r}
syndergaard.velo = graph.pitch.velo(syndergaard)
syndergaard.velo
```

#### Blake Treinen

```{r}
treinen.velo = graph.pitch.velo(treinen)
treinen.velo
```

#### Felipe Vazquez

```{r}
vazquez.velo = graph.pitch.velo(vazquez)
vazquez.velo
```

#### Justin Verlander

```{r}
verlander.velo = graph.pitch.velo(verlander)
verlander.velo
```

### Spin Rate {.tabset .tabset-fade}

Let's create our graping function.

```{r pitch_spin}
graph.pitch.spin = function(player) {
    graph = ggplot(player) +
        geom_step(aes(x = 1:length(player$spin),
                      y = player$spin,
                      color = player$pitch),
                  direction = "vh") +
        xlab("Pitches Thrown") + ylab("Spin Rate") + labs(color = "Pitch Type") +
        ggtitle(paste(player$name[1], player$year[1], "Spin Rate Chart", sep = " ")) +
        theme_minimal()
}
```

#### Patrick Corbin

```{r}
corbin.spin = graph.pitch.spin(corbin)
corbin.spin
```

#### Marcus Stroman

```{r}
stroman.spin = graph.pitch.spin(stroman)
stroman.spin
```

#### Noah Syndergaard

```{r}
syndergaard.spin = graph.pitch.spin(syndergaard)
syndergaard.spin
```

#### Blake Treinen

```{r}
treinen.spin = graph.pitch.spin(treinen)
treinen.spin
```

#### Felipe Vazquez

```{r}
vazquez.spin = graph.pitch.spin(vazquez)
vazquez.spin
```

#### Justin Verlander

```{r}
verlander.spin = graph.pitch.spin(verlander)
verlander.spin
```

# Analysis {.tabset .tabset-fade}

I'll be looking at a few specific Pittsburgh Pirates pitchers and looking at them from year to year.

## Jameson Taillon {.tabset .tabset-fade}

### Data Acquisition

Let's first read in our data for Taillon.

```{r}
taillon.data.2018 = scrape.data("2018-01-01", 592791)
taillon.data.2019 = scrape.data("2019-01-01", 592791)
taillon.2018 = filter.data(taillon.data.2018)
taillon.2019 = filter.data(taillon.data.2019)
```

### 2018

Now, let's just get some averages of Taillon's pitches for 2018.

```{r}
taillon.ff.2018 = taillon.2018[taillon.2018$pitch == "FF",]
taillon.ft.2018 = taillon.2018[taillon.2018$pitch == "FT",]
taillon.sl.2018 = taillon.2018[taillon.2018$pitch == "SL",]
taillon.cu.2018 = taillon.2018[taillon.2018$pitch == "CU",]
taillon.ch.2018 = taillon.2018[taillon.2018$pitch == "CH",]

taillon.ff.2018 = taillon.ff.2018[complete.cases(taillon.ff.2018),]
taillon.ft.2018 = taillon.ft.2018[complete.cases(taillon.ft.2018),]
taillon.sl.2018 = taillon.sl.2018[complete.cases(taillon.sl.2018),]
taillon.cu.2018 = taillon.cu.2018[complete.cases(taillon.cu.2018),]
taillon.ch.2018 = taillon.ch.2018[complete.cases(taillon.ch.2018),]
```

| Pitch | Average Velocity | Standard Deviation of Velocity | Average Spin Rate | Standard Deviation of Spin Rate |
| ----- | ---------------- | ------------------------------ | ----------------- | --------------------------------|
| 4-Seam Fastball | `r mean(taillon.ff.2018$velo)` | `r sd(taillon.ff.2018$velo)` | `r mean(taillon.ff.2018$spin)` | `r sd(taillon.ff.2018$spin)` |
| 2-Seam Fastball | `r mean(taillon.ft.2018$velo)` | `r sd(taillon.ft.2018$velo)` | `r mean(taillon.ft.2018$spin)` | `r sd(taillon.ft.2018$spin)` |
| Slider | `r mean(taillon.sl.2018$velo)` | `r sd(taillon.sl.2018$velo)` | `r mean(taillon.sl.2018$spin)` | `r sd(taillon.sl.2018$spin)` |
| Curveball | `r mean(taillon.cu.2018$velo)` | `r sd(taillon.cu.2018$velo)` | `r mean(taillon.cu.2018$spin)` | `r sd(taillon.cu.2018$spin)` |
| Changeup | `r mean(taillon.ch.2018$velo)` | `r sd(taillon.ch.2018$velo)` | `r mean(taillon.ch.2018$spin)` | `r sd(taillon.ch.2018$spin)` |

Now let's make some graphs.

```{r}
taillon.heatmap.out.2018 = graph.pitch.heatmap.out(taillon.2018)
taillon.heatmap.out.2018

taillon.heatmap.type.2018 = graph.pitch.heatmap.type(taillon.2018)
taillon.heatmap.type.2018

taillon.heatmap.velo.2018 = graph.pitch.heatmap.velo(taillon.2018)
taillon.heatmap.velo.2018

taillon.spin.2018 = graph.pitch.spin(taillon.2018)
taillon.spin.2018

taillon.velo.2018 = graph.pitch.velo(taillon.2018)
taillon.velo.2018

taillon.xmove.2018 = graph.pitch.xmovement(taillon.2018)
taillon.xmove.2018

taillon.ymove.2018 = graph.pitch.ymovement(taillon.2018)
taillon.ymove.2018
```

What's also important is to determine how many of his pitches were barrelled (strong contact).

```{r}
count.barrels = function(player) {
    s = sum(player$contact == 6)
    s
}

barrel.probability = function(player) {
    b = count.barrels(player)
    t = NROW(player$contact)
    p = (b * 1.0)/t
    p
}
```

Now let's take a look at his barrel probability for all his pitches.

| Pitch | Number of Barrels | Barrel Probability |
| ----- | ----------------- | ------------------ |
| All | `r count.barrels(taillon.2018)` | `r barrel.probability(taillon.2018)` |
| 4-Seam Fastball | `r count.barrels(taillon.ff.2018)` | `r barrel.probability(taillon.ff.2018)` |
| 2-Seam Fastball | `r count.barrels(taillon.ft.2018)` | `r barrel.probability(taillon.ft.2018)` |
| Slider | `r count.barrels(taillon.sl.2018)` | `r barrel.probability(taillon.sl.2018)` |
| Curveball | `r count.barrels(taillon.cu.2018)` | `r barrel.probability(taillon.cu.2018)` |
| Changeup | `r count.barrels(taillon.ch.2018)` | `r barrel.probability(taillon.ch.2018)` |

Let's also take a look at his pitch frequencies.

```{r}
pitch.count = function(player, type) {
    c = sum(player$pitch == type)
    c
}

pitch.frequency = function(player, type) {
    c = pitch.count(player, type)
    t = NROW(player$pitch)
    f = (c * 1.0) / t
    f
}
```

Let's view all his frequencies together in a tabular format.

| Pitch | Pitch Count | Pitch Frequency |
| ----- | ----------- | --------------- |
| 4-Seam Fastball | `r pitch.count(taillon.2018, "FF")` | `r pitch.frequency(taillon.2018, "FF")` |
| 2-Seam Fastball | `r pitch.count(taillon.2018, "FT")` | `r pitch.frequency(taillon.2018, "FT")` |
| Slider | `r pitch.count(taillon.2018, "SL")` | `r pitch.frequency(taillon.2018, "SL")` |
| Curveball | `r pitch.count(taillon.2018, "CU")` | `r pitch.frequency(taillon.2018, "CU")` |
| Changeup | `r pitch.count(taillon.2018, "CH")` | `r pitch.frequency(taillon.2018, "CH")` |

### 2019

Now, let's just get some averages of Taillon's pitches for 2019.

```{r}
taillon.ff.2019 = taillon.2019[taillon.2019$pitch == "FF",]
taillon.ft.2019 = taillon.2019[taillon.2019$pitch == "FT",]
taillon.sl.2019 = taillon.2019[taillon.2019$pitch == "SL",]
taillon.cu.2019 = taillon.2019[taillon.2019$pitch == "CU",]
taillon.ch.2019 = taillon.2019[taillon.2019$pitch == "CH",]

taillon.ff.2019 = taillon.ff.2019[complete.cases(taillon.ff.2019),]
taillon.ft.2019 = taillon.ft.2019[complete.cases(taillon.ft.2019),]
taillon.sl.2019 = taillon.sl.2019[complete.cases(taillon.sl.2019),]
taillon.cu.2019 = taillon.cu.2019[complete.cases(taillon.cu.2019),]
taillon.ch.2019 = taillon.ch.2019[complete.cases(taillon.ch.2019),]
```

| Pitch | Average Velocity | Standard Deviation of Velocity | Average Spin Rate | Standard Deviation of Spin Rate |
| ----- | ---------------- | ------------------------------ | ----------------- | --------------------------------|
| 4-Seam Fastball | `r mean(taillon.ff.2019$velo)` | `r sd(taillon.ff.2019$velo)` | `r mean(taillon.ff.2019$spin)` | `r sd(taillon.ff.2019$spin)` |
| 2-Seam Fastball | `r mean(taillon.ft.2019$velo)` | `r sd(taillon.ft.2019$velo)` | `r mean(taillon.ft.2019$spin)` | `r sd(taillon.ft.2019$spin)` |
| Slider | `r mean(taillon.sl.2019$velo)` | `r sd(taillon.sl.2019$velo)` | `r mean(taillon.sl.2019$spin)` | `r sd(taillon.sl.2019$spin)` |
| Curveball | `r mean(taillon.cu.2019$velo)` | `r sd(taillon.cu.2019$velo)` | `r mean(taillon.cu.2019$spin)` | `r sd(taillon.cu.2019$spin)` |
| Changeup | `r mean(taillon.ch.2019$velo)` | `r sd(taillon.ch.2019$velo)` | `r mean(taillon.ch.2019$spin)` | `r sd(taillon.ch.2019$spin)` |

Now let's make some graphs.

```{r}
taillon.heatmap.out.2019 = graph.pitch.heatmap.out(taillon.2019)
taillon.heatmap.out.2019

taillon.heatmap.type.2019 = graph.pitch.heatmap.type(taillon.2019)
taillon.heatmap.type.2019

taillon.heatmap.velo.2019 = graph.pitch.heatmap.velo(taillon.2019)
taillon.heatmap.velo.2019

taillon.spin.2019 = graph.pitch.spin(taillon.2019)
taillon.spin.2019

taillon.velo.2019 = graph.pitch.velo(taillon.2019)
taillon.velo.2019

taillon.xmove.2019 = graph.pitch.xmovement(taillon.2019)
taillon.xmove.2019

taillon.ymove.2019 = graph.pitch.ymovement(taillon.2019)
taillon.ymove.2019
```

What's also important is to determine how many of his pitches were barrelled (strong contact).

Now let's take a look at his barrel probability for all his pitches.

| Pitch | Number of Barrels | Barrel Probability |
| ----- | ----------------- | ------------------ |
| All | `r count.barrels(taillon.2019)` | `r barrel.probability(taillon.2019)` |
| 4-Seam Fastball | `r count.barrels(taillon.ff.2019)` | `r barrel.probability(taillon.ff.2019)` |
| 2-Seam Fastball | `r count.barrels(taillon.ft.2019)` | `r barrel.probability(taillon.ft.2019)` |
| Slider | `r count.barrels(taillon.sl.2019)` | `r barrel.probability(taillon.sl.2019)` |
| Curveball | `r count.barrels(taillon.cu.2019)` | `r barrel.probability(taillon.cu.2019)` |
| Changeup | `r count.barrels(taillon.ch.2019)` | `r barrel.probability(taillon.ch.2019)` |

Let's view all his frequencies together in a tabular format.

| Pitch | Pitch Count | Pitch Frequency |
| ----- | ----------- | --------------- |
| 4-Seam Fastball | `r pitch.count(taillon.2018, "FF")` | `r pitch.frequency(taillon.2018, "FF")` |
| 2-Seam Fastball | `r pitch.count(taillon.2018, "FT")` | `r pitch.frequency(taillon.2018, "FT")` |
| Slider | `r pitch.count(taillon.2018, "SL")` | `r pitch.frequency(taillon.2018, "SL")` |
| Curveball | `r pitch.count(taillon.2018, "CU")` | `r pitch.frequency(taillon.2018, "CU")` |
| Changeup | `r pitch.count(taillon.2018, "CH")` | `r pitch.frequency(taillon.2018, "CH")` |

## Richard Rodriguez

```{r}
rodriguez.data = scrape.data("2018-01-01", 593144)
rodriguez = filter.data(rodriguez.data)
```

Now, let's just get some averages of Rodriguez's pitches.

```{r}
rodriguez.ff = rodriguez[rodriguez$pitch == "FF",]
rodriguez.sl = rodriguez[rodriguez$pitch == "SL",]

rodriguez.ff = rodriguez.ff[complete.cases(rodriguez.ff),]
rodriguez.sl = rodriguez.sl[complete.cases(rodriguez.sl),]
```

| Pitch | Average Velocity | Standard Deviation of Velocity | Average Spin Rate | Standard Deviation of Spin Rate |
| ----- | ---------------- | ------------------------------ | ----------------- | --------------------------------|
| 4-Seam Fastball | `r mean(rodriguez.ff$velo)` | `r sd(rodriguez.ff$velo)` | `r mean(rodriguez.ff$spin)` | `r sd(rodriguez.ff$spin)` |
| Slider | `r mean(rodriguez.sl$velo)` | `r sd(rodriguez.sl$velo)` | `r mean(rodriguez.sl$spin)` | `r sd(rodriguez.sl$spin)` |

Now let's make some graphs.

```{r}
rodriguez.heatmap.out = graph.pitch.heatmap.out(rodriguez)
rodriguez.heatmap.out

rodriguez.heatmap.type = graph.pitch.heatmap.type(rodriguez)
rodriguez.heatmap.type

rodriguez.heatmap.velo = graph.pitch.heatmap.velo(rodriguez)
rodriguez.heatmap.velo

rodriguez.spin = graph.pitch.spin(rodriguez)
rodriguez.spin

rodriguez.velo = graph.pitch.velo(rodriguez)
rodriguez.velo

rodriguez.xmove = graph.pitch.xmovement(rodriguez)
rodriguez.xmove

rodriguez.ymove = graph.pitch.ymovement(rodriguez)
rodriguez.ymove
```

## Joe Musgrove

```{r}
musgrove.data = scrape.data("2018-01-01", 605397)
musgrove = filter.data(musgrove.data)
```

## Jordan Lyles

```{r}
lyles.data = scrape.data("2018-01-01", 543475)
lyles = filter.data(lyles.data)
```

## Kyle Crick

```{r}
crick.data = scrape.data("2018-01-01", 605195)
crick = filter.data(crick.data)
```